---
title: "poison_survey"
author: "Kar"
date: '2022-05-05'
output: 
  github_document: 
    toc: true
    toc_depth: 4
always_allow_html: yes
---


## SUMMARY

## R PACKAGES

```{r}
library(tidyverse)
library(skimr)
library(factoextra)
library(FactoMineR)

```


## INTRODUCTION

This project will analyse a survey dataset using one of the principal component methods. The technique of "Principal component methods" is known as one of the machine learning methods, belonging to the "unsupervised" branch, which can quickly summarise a dataset by revealing the most important variables (columns) or observations (rows) in the dataset. 

The data of this survey dataset was collected from students in a primary school who suffered from food poisoning. The dataset contains various type of information such as age, time, whether the student is sicked, sex, symptoms of food poisoning (nausea, vomiting, abdominal, fever, and diarrhea) and food they ate (potato, fish, mayo, courgette, cheese, ice cream).





## DATA PREPARATION

The dataset of this project is called "poison", which is a public dataset from the R package "FactoMineR".

### Data Import

Following code download the dataset from the package.

```{r}
data(poison)

```

Randomly sample 10 rows within the dataset:

```{r}
sample_n(poison, 10)

```

### Data Exploration

The dataset has 55 rows (55 students) and 15 columns recording the various type of variables (also known as feature).

Following code also shows the variable type allocated by R to each of the column. 

```{r}
glimpse(poison)

```

Factor "fct" is a variable type we usually allocate to variables that have categorisation nature. There is usually a conversion step required to perform by analyst to convert character "chr" variable such as "Yes" and "No" into factor "fct". This step seems like has been completed.

Insights from following summary, 

* Respondents have age between 4 to 88, with majority of them fall between 6 to 10 years old with a median of 8. Mean of 16.93 shouldn't be used for reference because 88 is a outlier.   
* Time should be time of recording the data.   
* The rest of the variables have two categories of either "Yes" or "No".  

```{r}
summary(poison)

```
Following is an alternative to see the structure of the dataset. Other than Age and time, the rest of the variables are factor type with 2 levels, of either "Y" or "No", which stand for "Yes" or "No". 

```{r}
str(poison)

```

Checking missing data in the dataset. 

```{r}
skim_without_charts(poison)

```


## EDA

### Missing 



## PRINCIPAL COMPONENT METHODS

## REFERENCE

KASSAMBARA A 2017, *Practical Guide To Principal Component Methods in R*, Edition 1, sthda.com